Suboptimality Bounds for Approximate State Abstraction

### 2016-02-01 paper/paper.tex
Some minor changes -- defns now are automatically counted

\documentclass{article}
\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure} 
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{algorithmic}
\usepackage{subfigure}
\usepackage{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}} % Fixes misbehavior of conflicting packages.
\usepackage{icml2016/icml2016} 
\usepackage[nolist]{acronym} % Acronym package -- give argument <dua> to suppress acronym usage

@@

% --- Notation Commands ---
\newcommand{\ep}{\widetilde \phi}
\newcommand{\epQ}{\ep_{Q^*}}
\newcommand{\epM}{\ep_{\text{model}}}
\newcommand\defn[1]{{\bf Definition:} #1}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\mcS}{\mathcal{S}}
\newcommand{\mcT}{\mathcal{T}}
\newcommand{\mcR}{\mathcal{R}}
\newcommand{\pM}{\phi_{\text{model}}}

@@

% Abstraction is a good idea b/c MDP has poly comp comp and samp comp in S, but S grows fast.
Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space~\cite{papadimitriou1987complexity,littman1995complexity}. Similarly, many \ac{RL} algorithms for solving \acp{MDP} are known to require a number of samples polynomial in the size of the state space~\cite{strehl2009reinforcement}. Although polynomial runtime or sample complexity may seem like a reasonable constraint, due to Bellman's curse of dimensionality, the size of the state space of an \ac{MDP} grows super-polynomially with the number of variables that characterize the domain. Thus, bounds polynomial in state space size often translate to intractable problems for sufficiently complicated tasks. \dnote{Would prefer a different example} For instance, a robot involved in a pick and place task might be able to employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider to do so grows exponentially with the number of objects it plans over CITE DABE ICAPS.

@@

\enote{Define $V_G$ and $V_A$ (value function in ground and value function in abstract)}
\enote{Say more about the weighting (that it sums to 1 but that's all we assume)}

@@

\defn{A $Q^*$-irrelevance abstraction, $\phi_{Q^*}$, is one such that:
\begin{equation}
\phi_{Q^*}(s_1) = \phi_{Q^*}(s_2) \rightarrow \forall a\ \left(Q^*(s_1,a) = Q^*(s_2,a)\right).
\end{equation}
}

\defn{A model-irrelevance abstraction, $\phi_{model}$, is one such that:
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \forall a\ \mathcal{R}_G(s_1,a) = \mathcal{R}_G(s_2,a). \\
\end{multline}
and
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \\ \forall a \sum_{s' \in \phi^{-1}(\phi(s_1))} \mathcal{T}_G(s_1,a,s') =\sum_{s' \in \phi^{-1}(\phi(s_2))} \mathcal{T}_G(s_2,a,s'). \\
\end{multline}
}

@@

% Theorem 1
{\bf Theorem 1:} There exist at least four classes of approximate state-aggregation functions, $\ep_{Q^*}$, $\ep_{\text{model}}$, $\ep_{\text{bolt}}$ and $\ep_{\text{mult}}$ for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\epsilon$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq poly(\epsilon).
\end{equation}

@@

% zzz wy not \begin{lemma} ? you know, to get the number right automatically?
{\bf Lemma 1:} When $\ep_{Q^*}$ is used to create the abstract state space $\mathcal{S}_A$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon}{(1-\gamma)^2}.
\end{equation}

@@

%Claim 1 zzz \begin{claim} ?
\textbf{Claim 1 -- Q-Values in the abstract \ac{MDP} closely resemble those in the ground \ac{MDP}:}
\begin{equation}
\label{eq:Q*Claim1}
\forall_{s_G \in \mathcal{S}_G, a} |Q_G(s_G, a) - Q_A(\epQ(s_G), a)| \leq \frac{\epsilon}{1-\gamma}.
\end{equation}

@@

%Claim 2
\textbf{Claim 2: the optimal action in the abstract MDP has a Q value in the ground which is nearly optimal:}

@@

%Claim 3
\textbf{Claim 3: we conclude Lemma 1 by Claim 2}

@@

{\bf Lemma 2:} When $S_A$ is created using a function of the $\ep_{model}$ type:
\begin{equation}
\label{eq:lemma2}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon + 2\gamma((|\mcS_G|-1)\epsilon)}{(1-\gamma)^3}
\end{equation}

{\bf Proof:} Our proof proceeds as follows: we demonstrate that states grouped aggregated under $\epM$ have Q values all within a particular bound. Consequently $\epM$ abstractions are interpretable as an instantiation of $\epQ$ with a particular $\epsilon$ -- understanding $\epM$ abstractions allows us to conclude our bound by applying Lemma 1.

@@

Since the Q values of ground states grouped under $\epM$ are strictly less than $B$, we can understand $\epM$ as a type of $\epQ$ with $\epsilon = B$. Applying Lemma 1 yields Equation \ref{eq:lemma2}.

>>>>

\documentclass{article}
\usepackage{times}
\usepackage{graphicx}
\usepackage{subfigure} 
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amsthm}%For stating numbered theorems
\usepackage{amsfonts}
\usepackage{algorithmic}
\usepackage{subfigure}
\usepackage{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}} % Fixes misbehavior of conflicting packages.
\usepackage{icml2016/icml2016} 
\usepackage[nolist]{acronym} % Acronym package -- give argument <dua> to suppress acronym usage

%THeorem declarations
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem{clm}{Claim}
\newtheorem{defn}{Definition}

@@

% --- Notation Commands ---
\newcommand{\ep}{\widetilde \phi}
\newcommand{\epQ}{\ep_{Q^*}}
\newcommand{\epM}{\ep_{\text{model}}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\mcS}{\mathcal{S}}
\newcommand{\mcT}{\mathcal{T}}
\newcommand{\mcR}{\mathcal{R}}
\newcommand{\pM}{\phi_{\text{model}}}

@@

% Abstraction is a good idea b/c MDP has poly comp comp and samp comp in S, but S grows fast.
Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space~\cite{papadimitriou1987complexity,littman1995complexity}. Similarly, many \ac{RL} algorithms for solving \acp{MDP} are known to require a number of samples polynomial in the size of the state space~\cite{strehl2009reinforcement}. Although polynomial runtime or sample complexity may seem like a reasonable constraint, due to Bellman's curse of dimensionality, the size of the state space of an \ac{MDP} grows super-polynomially with the number of variables that characterize the domain. Thus, bounds polynomial in state space size often translate to intractable problems for sufficiently complicated tasks. \dnote{Would prefer a different example} For instance, a robot involved in a pick and place task might be able to employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider to do so grows exponentially with the number of objects it plans over \enote{Cite Dabe Icaps}.

@@

\enote{Define $V_G$ and $V_A$ (value function in ground and value function in abstract)}
\enote{Say more about the weighting, w(s), (that it sums to 1 but that's all we assume)}

@@

\begin{defn}
A $Q^*$-irrelevance abstraction, $\phi_{Q^*}$, is one such that:
\begin{equation}
\phi_{Q^*}(s_1) = \phi_{Q^*}(s_2) \rightarrow \forall a\ \left(Q^*(s_1,a) = Q^*(s_2,a)\right).
\end{equation}
\end{defn}

\begin{defn}
A model-irrelevance abstraction, $\phi_{model}$, is one such that:
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \forall a\ \mathcal{R}_G(s_1,a) = \mathcal{R}_G(s_2,a). \\
\end{multline}
and
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \\ \forall a \sum_{s' \in \phi^{-1}(\phi(s_1))} \mathcal{T}_G(s_1,a,s') =\sum_{s' \in \phi^{-1}(\phi(s_2))} \mathcal{T}_G(s_2,a,s'). \\
\end{multline}
\end{defn}

@@

% Theorem 1
\begin{thm}
There exist at least four classes of approximate state-aggregation functions, $\ep_{Q^*}$, $\ep_{\text{model}}$, $\ep_{\text{bolt}}$ and $\ep_{\text{mult}}$ for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\epsilon$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq poly(\epsilon).
\end{equation}

\end{thm}

@@

\begin{lma}
\label{lma:Q*}
When $\ep_{Q^*}$ is used to create the abstract state space $\mathcal{S}_A$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon}{(1-\gamma)^2}.
\end{equation}
\end{lma}

@@

\begin{clm}
Q-Values in the abstract \ac{MDP} closely resemble those in the ground \ac{MDP}:
\begin{equation}
\label{eq:Q*Claim1}
\forall_{s_G \in \mathcal{S}_G, a} |Q_G(s_G, a) - Q_A(\epQ(s_G), a)| \leq \frac{\epsilon}{1-\gamma}.
\end{equation}
\end{clm}

@@

%Claim 2
\begin{clm}
\label{clm:optAbsActionNearOptGround}
The optimal action in the abstract MDP has a Q value in the ground which is nearly optimal.
\end{clm}

@@

%Claim 3
\begin{clm}
We conclude Lemma \ref{lma:Q*} by Claim \ref{clm:optAbsActionNearOptGround}.
\end{clm}

@@

%Model Lemma
\begin{lma}
\label{lma:model}
When $S_A$ is created using a function of the $\ep_{model}$ type:
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon + 2\gamma((|\mcS_G|-1)\epsilon)}{(1-\gamma)^3}
\end{equation}
\end{lma}

{\bf Proof:} Our proof proceeds as follows: we demonstrate that states grouped aggregated under $\epM$ have Q values all within a particular bound. Consequently $\epM$ abstractions are interpretable as an instantiation of $\epQ$ with a particular $\epsilon$ -- understanding $\epM$ abstractions allows us to conclude our bound by applying Lemma \ref{lma:Q*}.

@@

Since the Q values of ground states grouped under $\epM$ are strictly less than $B$, we can understand $\epM$ as a type of $\epQ$ with $\epsilon = B$. Applying Lemma \ref{lma:Q*} yields Lemma \ref{lma:model}.
